---
title: "Final"
author: "Samuel Lane & Nolan Scott"
format: 
  dashboard: 
    theme: sandstone
embed-resources: true
server: shiny
---

```{r}
#| warning: false
#| message: false
#| context: setup
library(tidyverse)
library(fmsb)
library(shiny)

climbs = read_csv("Climbs.csv", col_select = c(1, 2, 3, 4, 5, 6, 8, 11, 14, 15, 
                                               16, 17, 26, 28, 32, 34), 
na = c("NA", "", "N/A", "Don't know", "?", "??", "???", "Na", "unknown", 
       "Not sure", "no idea", "Unsure", "Not sure", "Nil", "Dont know",
       "don't know", "Dunno", "????", "No clue", "i suck!", "N/a", 
       "oooh...not sure", "ha! no need to train that yet", 
       "Never tried this", "Not calculated before", "n/a", "-", 
       " I don't know", "idk", "I don't climb routes", "I don't boulder"))


colnames(climbs) = c("time", "sex", "height", "weight", "arm_span", 
"time_climbed", "max_grade_boulder", "max_grade_sport", "frequency_climbing", 
"avg_hours_climbing_week", "avg_hours_training_week", "frequency_hangboarding", 
"frequency_endurance", "frequency_strength", "max_pullup", "max_pushup")

climbs = climbs |> mutate(
  height = as.numeric(parse_number(height)),
  weight = as.numeric(parse_number(weight)),
  arm_span  = as.numeric(parse_number(arm_span)),
  max_grade_boulder  = as.integer(parse_number(max_grade_boulder)),
  max_grade_sport  = as.integer(max_grade_sport),
  frequency_climbing  = as.integer(frequency_climbing),
  avg_hours_climbing_week  = as.integer(avg_hours_climbing_week),
  avg_hours_training_week  = as.integer(avg_hours_training_week),
  frequency_hangboarding  = as.integer(frequency_hangboarding),
  frequency_endurance  = as.integer(frequency_endurance),
  frequency_strength  = as.integer(frequency_strength),
  max_pullup  = as.integer(parse_number(max_pullup)),
  max_pushup  = as.integer(parse_number(max_pushup))
)


training_vars <- c("frequency_climbing", "avg_hours_climbing_week", 
                   "avg_hours_training_week", "frequency_hangboarding",
                   "frequency_endurance", "frequency_strength",
                   "max_pullup", "max_pushup")

pretty_names <- c(
  "Sessions/week", "Hours climbing", "Hours training",
  "Hangboard freq", "Endurance freq", "Strength freq",
  "Max pull-ups", "Max push-ups")
```

```{css}
.cell.bslib-grid-item.html-fill-item {
  height: 0 !important;
  min-height: 0 !important;
  flex: 0 0 0 !important;
  overflow: hidden !important;
}

.card-body{
  font-size: 20px;
}
```

# Introduction

::: {.card}
We chose the Climbs.csv dataset we found online as we are both avid climbers. We had to clean the data very intensively as not many people answered the questions in the same formatting. This included removing almost half of the columns that were so bad they could not be saved. What we ended up with was very useful and exciting data. 

Now with clean data we wanted to explore many things such as:
<ul>
  <li>Ape index vs Max V Grade</li>
  <li>Correlation of training on Max V Grade</li>
  <li>BMI within V grades</li>
</ul>

We ended up mainly talk about Max V Grade instead of sport grade because there were way more boulderers who took the survey.
:::
# Results 1

```{r}
#| warning: false
#Had to look up code for IQR to remove outliers
climbsAH = climbs |> filter(
  arm_span >= quantile(arm_span, 0.25, na.rm = TRUE) - 1.5 * IQR(arm_span, na.rm = TRUE), 
  arm_span <= quantile(arm_span, 0.75, na.rm = TRUE) + 1.5 * IQR(arm_span, na.rm = TRUE),
  height >= quantile(height, 0.25, na.rm = TRUE) - 1.5 * IQR(height, na.rm = TRUE), 
  height <= quantile(height, 0.75, na.rm = TRUE) + 1.5 * IQR(height, na.rm = TRUE)
)
#arm span vs max grade bouldering
ggplot(data = climbsAH, aes(x = arm_span, y = max_grade_boulder), na.rm = TRUE) + 
  geom_smooth()
#ape index vs max grade bouldering
ggplot(data = climbsAH, aes(x = arm_span - height, y = max_grade_boulder)) + 
  geom_smooth()
```



# Results 2

```{r}
#| warning: false

climbsP = climbs |> filter(
  max_pullup >= quantile(max_pullup, 0.25, na.rm = TRUE) - 1.5 * IQR(max_pullup, na.rm = TRUE), 
  max_pullup <= quantile(max_pullup, 0.75, na.rm = TRUE) + 1.5 * IQR(max_pullup, na.rm = TRUE),
  max_pushup >= quantile(max_pushup, 0.25, na.rm = TRUE) - 1.5 * IQR(max_pushup, na.rm = TRUE), 
  max_pushup <= quantile(max_pushup, 0.75, na.rm = TRUE) + 1.5 * IQR(max_pushup, na.rm = TRUE)
)

#pullup + pushup vs max grade
ggplot(data = climbsP, na.rm = TRUE) + 
  geom_smooth(aes(x = max_pullup, y = max_grade_boulder)) +
  geom_smooth(aes(x = max_pushup, y = max_grade_boulder))
                

  ggplot(data = climbsP, na.rm = TRUE) + 
  geom_smooth(aes(x = max_pullup, y = max_grade_boulder)) +
  geom_smooth(aes(x = max_pushup, y = max_grade_boulder)) +
  xlim(0, 35)


```

```{r}

# Compute correlations of training variables with max_grade_boulder
training_vars <- c("frequency_climbing", "avg_hours_climbing_week", 
                   "avg_hours_training_week", "frequency_hangboarding",
                   "frequency_endurance", "frequency_strength",
                   "max_pullup", "max_pushup")

cors <- climbs |>
  select(max_grade_boulder, all_of(training_vars)) |>
  drop_na() |>
  summarise(across(all_of(training_vars),
                   ~cor(.x, max_grade_boulder, use = "complete.obs"))) |>
  round(3)

radar_data <- bind_rows(
  setNames(rep(list(1),  length(training_vars)), training_vars),  # max
  setNames(rep(list(0),  length(training_vars)), training_vars),  # min (was -1)
  cors
)

colnames(radar_data) <- c(
  "Sessions\n/week", "Hours\nclimbing", "Hours\ntraining",
  "Hangboard\nfreq", "Endurance\nfreq", "Strength\nfreq",
  "Max\npull-ups", "Max\npush-ups"
)

par(mar = c(1, 1, 2, 1), bg = "white")

radarchart(
  radar_data,
  axistype   = 1,
  pcol       = "#2171b5",
  pfcol      = adjustcolor("#2171b5", alpha.f = 0.25),
  plwd       = 2.5,
  cglcol     = "grey75",
  cglty      = 1,
  cglwd      = 0.6,
  axislabcol = "grey40",
  vlcex      = 0.85,
  calcex     = 0.7,
  caxislabels = c("0", "0.25", "0.5", "0.75", "1"),  # updated
  title      = "Correlation with max boulder grade"
)
```

```{r}
#| context: setup
library(shiny)
library(fmsb)
library(tidyverse)

training_vars <- c("frequency_climbing", "avg_hours_climbing_week", 
                   "avg_hours_training_week", "frequency_hangboarding",
                   "frequency_endurance", "frequency_strength",
                   "max_pullup", "max_pushup")

pretty_names <- c(
  "Sessions/week", "Hours climbing", "Hours training",
  "Hangboard freq", "Endurance freq", "Strength freq",
  "Max pull-ups", "Max push-ups"
)
```

```{r}
#| content: card-sidebar

checkboxGroupInput(
  inputId = "training_vars",
  label = "Select variables:",
  choices = pretty_names,
  selected = pretty_names
)
```

```{r}
#| title: Training Variables vs Max Boulder Grade

plotOutput("trainingPlot")
```

```{r}
#| context: server

output$trainingPlot <- renderPlot({
  selected <- input$training_vars
  req(length(selected) >= 1)
  
  climbs |>
    select(max_grade_boulder, all_of(training_vars)) |>
    drop_na() |>
    setNames(c("max_grade_boulder", pretty_names)) |>
    pivot_longer(-max_grade_boulder, names_to = "Parameter", values_to = "Value") |>
    filter(Parameter %in% selected) |>
    ggplot(aes(x = Value, y = max_grade_boulder, color = Parameter)) +
    geom_smooth(se = FALSE) +
    facet_wrap(~Parameter, scales = "free_x") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(y = "Max Boulder Grade", x = "")
})
```





```{r}
#| warning: FALSE
climbsB = climbs |> filter(
  !is.na(weight), !is.na(height), !is.na(max_grade_boulder),
  weight >= quantile(weight, 0.25) - 1.5 * IQR(weight), 
  weight <= quantile(weight, 0.75) + 1.5 * IQR(weight),
  height >= quantile(height, 0.25) - 1.5 * IQR(height), 
  height <= quantile(height, 0.75) + 1.5 * IQR(height)
)
climbsB |>
  mutate(bmi = weight / (height / 100)^2) |>
  ggplot(aes(x = factor(max_grade_boulder), y = bmi)) +
  geom_violin(fill = "steelblue", alpha = 0.4) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
  labs(x = "Max Boulder Grade (V scale)", y = "BMI")
```


# Summary

Stuff goes here

