---
title: "Rock Climbing V Grades Study"
author: "Samuel Lane & Nolan Scott"
format: 
  dashboard: 
    theme: sandstone
embed-resources: true
server: shiny
---

```{r}
#| warning: false
#| message: false
#| context: setup
library(tidyverse)
library(fmsb)
library(shiny)

climbs = read_csv("Climbs.csv", col_select = c(1, 2, 3, 4, 5, 6, 8, 11, 14, 15, 
                                               16, 17, 26, 28, 32, 34), 
na = c("NA", "", "N/A", "Don't know", "?", "??", "???", "Na", "unknown", 
       "Not sure", "no idea", "Unsure", "Not sure", "Nil", "Dont know",
       "don't know", "Dunno", "????", "No clue", "i suck!", "N/a", 
       "oooh...not sure", "ha! no need to train that yet", 
       "Never tried this", "Not calculated before", "n/a", "-", 
       " I don't know", "idk", "I don't climb routes", "I don't boulder"))


colnames(climbs) = c("time", "sex", "height", "weight", "arm_span", 
"time_climbed", "max_grade_boulder", "max_grade_sport", "frequency_climbing", 
"avg_hours_climbing_week", "avg_hours_training_week", "frequency_hangboarding", 
"frequency_endurance", "frequency_strength", "max_pullup", "max_pushup")

climbs = climbs |> mutate(
  height = as.numeric(parse_number(height)),
  weight = as.numeric(parse_number(weight)),
  arm_span  = as.numeric(parse_number(arm_span)),
  max_grade_boulder  = as.integer(parse_number(max_grade_boulder)),
  max_grade_sport  = as.integer(max_grade_sport),
  frequency_climbing  = as.integer(frequency_climbing),
  avg_hours_climbing_week  = as.integer(avg_hours_climbing_week),
  avg_hours_training_week  = as.integer(avg_hours_training_week),
  frequency_hangboarding  = as.integer(frequency_hangboarding),
  frequency_endurance  = as.integer(frequency_endurance),
  frequency_strength  = as.integer(frequency_strength),
  max_pullup  = as.integer(parse_number(max_pullup)),
  max_pushup  = as.integer(parse_number(max_pushup))
)



cols <- c("weight", "height", "arm_span", "frequency_climbing", "avg_hours_climbing_week", "avg_hours_training_week", "frequency_hangboarding", "frequency_endurance", "frequency_strength", "max_pullup", "max_pushup")

climbs <- climbs |> filter(across(all_of(cols), ~
  .x >= quantile(.x, 0.25, na.rm = TRUE) - 1.5 * IQR(.x, na.rm = TRUE) &
  .x <= quantile(.x, 0.75, na.rm = TRUE) + 1.5 * IQR(.x, na.rm = TRUE)
))


training_vars <- c("frequency_climbing", "avg_hours_climbing_week", 
                   "avg_hours_training_week", "frequency_hangboarding",
                   "frequency_endurance", "frequency_strength",
                   "max_pullup", "max_pushup")

pretty_names <- c(
  "Sessions/week", "Hours climbing", "Hours training",
  "Hangboard freq", "Endurance freq", "Strength freq",
  "Max pull-ups", "Max push-ups")
```

```{css}
.cell.bslib-grid-item.html-fill-item {
  height: 0 !important;
  min-height: 0 !important;
  flex: 0 0 0 !important;
  overflow: hidden !important;
}

.card-body{
  font-size: 20px;
}
```

# Introduction

::: {.card}

What are the greatest contributing factors to someones maximum climbing ability?

We chose the Climbs.csv dataset we found online as we are both avid climbers. We had to clean the data very intensively as not many people answered the questions in the same formatting. This included removing almost half of the columns that were so bad they could not be saved. What we ended up with was very useful and exciting data. 

<strong>Now with clean data we wanted to explore many things such as:</strong>
<ul>
  <li>Ape index vs Max V Grade</li>
  <li>Correlation of training on Max V Grade</li>
  <li>BMI within V grades</li>
</ul>

We ended up mainly talk about Max V Grade instead of sport grade because there were way more boulderers who took the survey.

<strong>New skills learned:</strong>
<ul>
  <li>We learned how to get bullet points in the dashboard</li>
  <li>Found Violin plot for a cool way to show data</li>
  <li>Spider plot because it condensed our findings</li>
  <li>css to fix weird formatting in the dashboard</li>
  <li>Not new but we used all the skills learned in class</li>
</ul>
:::
# Results 1

```{r}
#| warning: false

#ape index vs max grade bouldering
ggplot(data = climbs, aes(x = arm_span - height, y = max_grade_boulder)) + 
  geom_smooth()
```

```{r}
#| warning: FALSE
climbsB = climbs |> filter(
  !is.na(weight), !is.na(height), !is.na(max_grade_boulder),
)
climbsB |>
  mutate(bmi = weight / (height / 100)^2) |>
  ggplot(aes(x = factor(max_grade_boulder), y = bmi)) +
  geom_violin(fill = "steelblue", alpha = 0.4) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.5) +
  labs(x = "Max Boulder Grade (V scale)", y = "BMI")
```

# Results 2




```{r}

# Compute correlations of training variables with max_grade_boulder
training_vars <- c("frequency_climbing", "avg_hours_climbing_week", 
                   "avg_hours_training_week", "frequency_hangboarding",
                   "frequency_endurance", "frequency_strength",
                   "max_pullup", "max_pushup")

cors <- climbs |>
  select(max_grade_boulder, all_of(training_vars)) |>
  drop_na() |>
  summarise(across(all_of(training_vars),
                   ~cor(.x, max_grade_boulder, use = "complete.obs"))) |>
  round(3)

radar_data <- bind_rows(
  setNames(rep(list(1),  length(training_vars)), training_vars),  # max
  setNames(rep(list(0),  length(training_vars)), training_vars),  # min (was -1)
  cors
)

colnames(radar_data) <- c(
  "Sessions\n/week", "Hours\nclimbing", "Hours\ntraining",
  "Hangboard\nfreq", "Endurance\nfreq", "Strength\nfreq",
  "Max\npull-ups", "Max\npush-ups"
)

par(mar = c(1, 1, 2, 1), bg = "white")

radarchart(
  radar_data,
  axistype   = 1,
  pcol       = "#2171b5",
  pfcol      = adjustcolor("#2171b5", alpha.f = 0.25),
  plwd       = 2.5,
  cglcol     = "grey75",
  cglty      = 1,
  cglwd      = 0.6,
  axislabcol = "grey40",
  vlcex      = 0.85,
  calcex     = 0.7,
  caxislabels = c("0", "0.25", "0.5", "0.75", "1"),  # updated
  title      = "Correlation with max boulder grade"
)
```

```{r}
#| context: setup
library(shiny)
library(fmsb)
library(tidyverse)

training_vars <- c("frequency_climbing", "avg_hours_climbing_week", 
                   "avg_hours_training_week", "frequency_hangboarding",
                   "frequency_endurance", "frequency_strength",
                   "max_pullup", "max_pushup")

pretty_names <- c(
  "Sessions/week", "Hours climbing", "Hours training",
  "Hangboard freq", "Endurance freq", "Strength freq",
  "Max pull-ups", "Max push-ups"
)



```

```{r}
#| content: card-sidebar

checkboxGroupInput(
  inputId = "training_vars",
  label = "Select variables:",
  choices = pretty_names,
  selected = pretty_names
)
```

```{r}
#| title: Training Variables vs Max Boulder Grade

plotOutput("trainingPlot")
```

```{r}
#| context: server

output$trainingPlot <- renderPlot({
  selected <- input$training_vars
  req(length(selected) >= 1)
  
  climbs |>
    select(max_grade_boulder, all_of(training_vars)) |>
    drop_na() |>
    setNames(c("max_grade_boulder", pretty_names)) |>
    pivot_longer(-max_grade_boulder, names_to = "Parameter", values_to = "Value") |>
    filter(Parameter %in% selected) |>
    ggplot(aes(x = Value, y = max_grade_boulder, color = Parameter)) +
    geom_smooth(se = FALSE) +
    facet_wrap(~Parameter, scales = "free_x") +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(y = "Max Boulder Grade", x = "")
})
```








# Summary

::: {.card}
Surprisingly there seemed to be no correlation between max boulder grade and a persons ape index, dispite many existing myths in the climbing community. So we turned to other predictors, BMI seemed promising as our violin graphs did show a tapering to a specific BMI of around 22-23 however we knew a better predictor existed. 

Looking at the training data, the best predictor is the max number of pull ups a person can do, followed by times climbing per week (associated with total hours climbing per week) and hang board frequency. Most surprisingly was the lack of positive correlation with endurancy and strength training.


This is a link to the dataset we used [Our Data :)]

[Our Data :)]: https://docs.google.com/spreadsheets/d/1J6d45EqIlIsIqNdi2X-Zl-EGFxf9d9T3R_W55xrpEAs/edit?gid=1650492946#gid=1650492946

If you would like to learn more, here are some additional resources that will be helpful

 - [Sport Climbing Predictors]
  - [Additional Pullup Data]
 - [Climber Data]

[Sport Climbing Predictors]: https://latticetraining.com/blog/what-is-the-number-1-measure-of-performance-for-sport-climbers
[Additional Pullup Data]: https://www.powercompanyclimbing.com/blog/heatmaps
[Climber Data]: https://www.kaggle.com/datasets/jordizar/climb-dataset
:::

